Preserving Reciprocal Consistency in Distributed Graph Databases

### 2020-04-10 paper/abstract.tex
camera-ready

\begin{abstract}
In our earlier work, we identified \emph{reciprocal consistency} as an important constraint specific to graph databases. If it can be lost even with a negligible small probability, subsequent inconsistent reads followed by writes can corrupt a distributed graph database within a time period extremely short relative to database lifetime. Reciprocal consistency can of course be maintained as a part of enforcing any known isolation guarantee incurring well established performance costs. However, in practice distributed graph databases are often built atop BASE databases with no isolation guarantees, profiting from increased performance but leaving them susceptible to rapid corruption. A lightweight concurrency control protocol ensuring reciprocal consistency is presented, catering for application programmers that are interested in maintaining performance and the structural integrity of their distributed graph database. Protocol performance is evaluated through simulations.
\end{abstract}

>>>>

\begin{abstract}
  Our earlier work identifies \emph{reciprocal consistency} as an important property that must be preserved in distributed graph databases.
  It also demonstrates that a failure to do so seriously undermines the integrity of the database itself in the long term.
  Reciprocal consistency can be maintained as a part of enforcing any known isolation guarantee and such an enforcement is also known to lead to reduction in performance. Therefore, in practice, distributed graph databases are often built atop BASE databases with no isolation guarantees, benefiting from good performance but leaving them susceptible to corruption due to violations of reciprocal consistency.
  This paper designs and presents a lightweight, locking-free protocol and then evaluates the protocol's abilities to preserve reciprocal consistency and also offer good throughput. Our evaluations establish that  the protocol can offer both integrity guarantees and sound performance when the value of its parameter is chosen appropriately.
\end{abstract}

### 2020-04-10 paper/appendix.tex
camera-ready

Under the \tDelta protocol each interleaving in Figure \ref{fig:conf} can be given a probability of occurring. Figures \ref{fig:a} and \ref{fig:b} are equivalent and can be summarized by the same probability. Letting $t_x = 0$, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( T_x >  \Delta + D) \cap (T_y > \Delta-D) \right]$$ The arrival times of $T_x,T_y$ are assumed exponentially distributed, $T \sim \exp (\rho)$.
Where, $ \rho  = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.
The probability of $d$ exceeding $\Delta$ is exponentially distributed with mean $\mu$, $D\sim \exp(\mu)$.
Therefore,
\begin{dmath*}
   q^{new}_{i} =  {P \left[ ( T_1 >  \Delta + D) \cap (T_2 > \Delta-D)  \right]} \\
   =  \int_{0}^{\Delta}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} d} e^{-\mu (\Delta+d)} e^{-\mu (\Delta-d)} dd + \int_{\Delta}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} d} e^{-\mu (\Delta+d)} dd  \\
   =  e^{-2 d \mu} - \left( \frac{\mu}{\frac{\lambda P_i}{2 N_i} + \mu} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\mu)d} \\
 \end{dmath*}
To simply the derivation of the conflict probability we assume FIFO which rules out the interleaving in Figure \ref{fig:c}.
 % To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
 % \begin{align*}
 %   P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
 %   1 - e^{- \delta \Delta} & = 1 - \epsilon \\
 %   e^{- \delta d} & = \epsilon  \\
 %   \Delta & = - \frac{\ln(\epsilon)}{\delta}
 % \end{align*}

>>>>

Edges are divided into $T$ types, $i= 1,2, ..., T$. The number of edges of type $i$ is represented by $n_i$. 
The probability an edge of type $i$ is accessed by a given operation is represented by $p_i$.  (Exact values taken by $n_i$ and $p_i$ are given in Section \ref{sec:evaluation}.)

The arrival time, $a$, of a transaction, $T$, is assumed exponentially distributed with rate $\rho$. 
Where $ \rho  = \frac{\lambda p_i}{2n_i}$ is the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$.

The time interval, $d$, that elapses between transaction $T$ completing its tentative write at one end of a distributed edge and starting at another end, is assumed exponentially distributed with rate $\mu$.

Consider the interleaving in Fig \ref{fig:a} and assume $T_x$ arrives at $S_i$ at time $0$. Then, $T_x$ arrives at $S_j$ after $d_x$. Assume $T_y$ arrives at $S_j$ at some time $a$. Then, $T_y$ arrives at $S_i$ at time $a  + d_y$.
Half-corruption occurs under the following conditions:
\begin{enumerate}[(i)]
    \item At $S_j$, $d_x > a + \Delta$
    \item At $S_i$, $a  + d_y > \Delta$
\end{enumerate}
The conflict probability, $q^{new}_{i}$, for edge type $i$ is given by,

\begin{dmath*}
   q^{new}_{i} =  {P \left[ (d_x > a + \Delta) \cap (d_y > \Delta - a)  \right]} \\
   =  \int_{0}^{\Delta}  \frac{\lambda p_i}{2 n_i} e^{-\frac{\lambda p_i}{2 n_i} a} e^{-\mu (\Delta+a)} e^{-\mu (\Delta-a)} da + \int_{\Delta}^{\infty} \frac{\lambda p_i}{2 n_i} e^{-\frac{\lambda p_i}{2 n_i} a} e^{-\mu (\Delta+a)} da \\
   =  e^{-2 a \mu} - \left( \frac{\mu}{\frac{\lambda p_i}{2 n_i} + \mu} \right) e^{-(\frac{\lambda p_i}{2 n_i}+ 2\mu)a} \\
 \end{dmath*}
%The interleaving in Fig \ref{fig:a} is equivalent to Fig \ref{fig:b} where $T_y$ arrives at $S_j$ at time $0$ and $S_i$ at $a$ while $T_x$ starts in $S_i$ at $a$ and reaches $S_j$ at $a+d_x$.

From the perspective of computing $q^{new}_{i}$, Fig \ref{fig:b} is equivalent to Fig \ref{fig:a} and the above expression holds.  
Fig \ref{fig:c} is ignored as it can be avoided by having all first tentative updates on a given edge at a given server (e.g., at $S_i$) be numbered sequentially and having the second tentative attempts  processed (at $S_j$) as per this sequence number.

%To simply the derivation of the conflict probability we assume FIFO which rules out the interleaving in Figure \ref{fig:c}.

 % To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
 % \begin{align*}
 %   P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
 %   1 - e^{- \delta \Delta} & = 1 - \epsilon \\
 %   e^{- \delta d} & = \epsilon  \\
 %   \Delta & = - \frac{\ln(\epsilon)}{\delta}
 % \end{align*}

### 2020-04-10 paper/figures/hc-edge.tex
section 2
camera-ready

\begin{figure}[H]
  \centering
  \begin{tikzpicture}[node distance=2cm]
    \node (v1) [big_vertex,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};

@@

    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year} = 1937}}} (2.4,0);
  \end{tikzpicture}
  \caption{A half-corrupted edge.}
  \label{half-corrupted}
\end{figure}

>>>>

\begin{tikzpicture}[node distance=2cm]
  \node (v1) [big_vertex,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};

@@

  \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year}:1937}}} (2.4,0);
\end{tikzpicture}

### 2020-04-10 paper/main.tex
more refactoring++
camera-ready

\usepackage{xspace}
\usepackage[labelformat=simple]{subcaption}
\usepackage{tikz}
\usepackage{fancyvrb}
\usepackage{xcolor}
\usepackage{breqn}

\setcopyright{acmcopyright}
\copyrightyear{2020}
\acmYear{2020}
\acmDOI{10.1145/1122445.1122456}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[PaPoC '20]{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data}{April 27, 2020}{Heraklion, Crete, Greece}
\acmBooktitle{PaPoC '20: Proceedings of the 7th Workshop on Principles and Practice of Consistency for Distributed Data, April 27, 2020, Heraklion, Crete, Greece}
\acmPrice{15.00}
% \acmISBN{978-1-4503-XXXX-X/18/06}

% Submission ID.
% Use this when submitting an article to a sponsored event. You'll
% receive a unique submission ID from the organizers
% of the event, and this ID should be used as the parameter to this command.
%\acmSubmissionID{123-A56-BU3}

@@

% \begin{CCSXML}
% <ccs2012>
%  <concept>
%   <concept_id>10010520.10010553.10010562</concept_id>
%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
%   <concept_significance>500</concept_significance>
%  </concept>
%  <concept>
%   <concept_id>10010520.10010575.10010755</concept_id>
%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
%   <concept_significance>300</concept_significance>
%  </concept>
%  <concept>
%   <concept_id>10010520.10010553.10010554</concept_id>
%   <concept_desc>Computer systems organization~Robotics</concept_desc>
%   <concept_significance>100</concept_significance>
%  </concept>
%  <concept>
%   <concept_id>10003033.10003083.10003095</concept_id>
%   <concept_desc>Networks~Network reliability</concept_desc>
%   <concept_significance>100</concept_significance>
%  </concept>
% </ccs2012>
% \end{CCSXML}

\ccsdesc[500]{Data Management~Graph Databases}
\ccsdesc[300]{Data Management~Reciprocal Consistency}
\ccsdesc[300]{Data Management~Concurrency Control}

\keywords{Graph Databases, Reciprocal Consistency, BASE}

@@

\input{introduction}

\input{reciprocal-consistency}

\input{dist-graphdb}

\input{corruption}

\input{delta}

\input{modeling}

\input{evaluation}

\input{conclusion}

% \appendix
% \section{Appendix}

% Figure \ref{corruption-again} presents an interleaving when a distributed edge can become half-corrupted under the \tDelta protocol.

% Transaction $T_x$ arrives at $S_i$ at time $t_x$ (let $t=0$) and writes tentatively, with the message delay between servers for $T_x$ to write the edge at $S_j$ being $M_x$. Transaction $T_y$ arrives at $S_j$ at time $t_y$ ($t_y > t_x$) and writes the distributed edge, $T_y$ then takes time $M_y$ to write the edge at $S_i$. If $T_x$ arrives at $S_j$ after $t_y + \Delta$ and $T_y$ arrives at $S_i$ after $t_x + \Delta$ the edge can become half-corrupted. Letting $t_x = 0 $, the probability that $T_x$ and $T_y$ conflict can be formulated as: $$ P \left[ ( T_x >  \Delta + X) \cap (T_y > \Delta-X) \right]$$.

% The arrival times of $T_x,T_y$ are assumed exponentially distributed , $T \sim \exp (\rho)$. Where, $ \rho  = \frac{\lambda P_i}{2N_i}$, the probability a given operation accesses the incorrect record of a half-corrupted edge of type $i$. The transmission times between servers $M_1, M_2 \sim \exp (\delta)$ and are \emph{iid}.

% % Therefore,  $$ g(y) = \delta e^{-\delta y} $$

% Therefore,
% \begin{dmath*}
%   q^{new}_{i} =  {P \left[ ( T_1 >  d + X) \cap (T_2 > d-X)  \right]} \\
%   =  \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} e^{-\delta (d-x)} dx + \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta (d+x)} dx  \\
%   =  e^{-2 d \delta} - \left( \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} \right) e^{-(\frac{\lambda P_i}{2 N_i}+ 2\delta)d} \\
% \end{dmath*}

% To choose to bound $\Delta$ consider the probability of the message delay exceeding  $\Delta$.
% \begin{align*}
%   P \left[ M > \Delta \right] & = 1 - e^{- \delta \Delta} \\
%   1 - e^{- \delta \Delta} & = 1 - \epsilon \\
%   e^{- \delta d} & = \epsilon  \\
%   \Delta & = - \frac{\ln(\epsilon)}{\delta}
% \end{align*}

% Substituting $\Delta$ into the above equation yields the conflict probability for a given $\epsilon$. For example, $ e = 0.001$ gives a $0.001$ \% probability the message delay exceeds $\Delta$, for this  $\Delta = 0.035s$
% \begin{align}
%   e^{- \lambda d} =  e^{-\delta d \frac{\lambda}{\delta}} = \epsilon^{\frac{\lambda}{\delta}} \label{eqn7}
% \end{align}
% Therefore, from (\ref{eqn5} and  (\ref{eqn7}),
% \begin{align}
%   & = \epsilon^2 - \frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \epsilon^2 \\
%   & = \epsilon^2 \left[1 -\frac{\delta}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-\frac{\lambda P_i}{2 N_i} d} \right]
% \end{align}

%Where, $ E \left[ T_1 \right] = E \left[ T_2\right] = \frac{1}{\delta}$.
%Therefore, $$ f(x) =\frac{\lambda P_i}{2N_i} e^{-\frac{\lambda P_i}{2N_i} x} $$

%  =   \int_{0}^{d}  \frac{\lambda P_i}{2 N_i} e^{-\frac{\lambda P_i}{2 N_i} x} e^{-\delta 2 d} dx +  \int_{d}^{\infty} \frac{\lambda P_i}{2 N_i} e^{-\delta d} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)x}  dx  \\
%         =   e^{-2 d \delta} \left[1 - e^{-\frac{\lambda P_i}{2 N_i} d} \right] + \frac{\frac{\lambda P_i}{2 N_i} e^{-\delta d}}{\frac{\lambda P_i}{2 N_i} + \delta} e^{-(\frac{\lambda P_i}{2 N_i} + \delta)d} \\

@@

>>>>

\usepackage{enumerate}
\usepackage{xspace}
\usepackage[labelformat=simple]{subcaption}
\usepackage{tikz}
\usepackage{fancyvrb}
\usepackage{xcolor}
\usepackage{breqn}

\input{commands}

\copyrightyear{2020}
\acmYear{2020}
\setcopyright{acmcopyright}
\acmConference[PaPoC '20]{7th Workshop on Principles and Practice of Consistency for Distributed Data}{April 27, 2020}{Heraklion, Greece}
\acmBooktitle{7th Workshop on Principles and Practice of Consistency for Distributed Data (PaPoC '20), April 27, 2020, Heraklion, Greece}
\acmPrice{15.00}
\acmDOI{10.1145/3380787.3393675}
\acmISBN{978-1-4503-7524-5/20/04}

@@

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10002951.10002952.10003190.10003193</concept_id>
       <concept_desc>Information systems~Database transaction processing</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10002951.10002952.10002953.10010146</concept_id>
       <concept_desc>Information systems~Graph-based database models</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Information systems~Database transaction processing}
\ccsdesc[500]{Information systems~Graph-based database models}

\keywords{Graph Databases, Concurrency Control, Reciprocal Consistency}

@@

\input{sec1-introduction}

\input{sec2-reciprocal-consistency}

\input{sec3-dist-graphdb}

\input{sec4-corruption}

\input{sec5-delta}

\input{sec6-modeling}

\input{sec7-evaluation}

\input{sec8-conclusion}

@@

\appendix

\input{appendix}

### 2020-04-10 paper/sec1-introduction.tex
camera-ready

In practice, graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node graph database and motivating the need for distributed graph databases. A common distributed graph database design pattern is to first partition graph data over several machines in a cluster; resulting in a number of \emph{distributed edges}, an edge's reciprocal pointers reside in different partitions. Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted that violations of reciprocal consistency in distributed edges introduce corruption into the database. Moreover, due to the \emph{Scale-Free} \cite{ScaleFree} property exhibited by many real world graphs, this corruption can propagate through the database at alarmingly rates.

When, for example, a BASE database \cite{Pritchett2008} is adapted with a graph processing layer, then violations of reciprocal consistency will occur if that adaptation provides no concurrency control for operations that span partitions in order to offer higher performance. This paper proposes a simple concurrency control protocol, called \tDelta protocol, that does not impede performance adversely. That is because the protocol is exclusively designed for one purpose only: reciprocal consistency in distributed edges.  Its design leverages the fact that the updating of end pointers of a given distributed edge immediately follow each other and the small interval between them is the window for possible conflicts between concurrent updates.

The paper is organized as follows: Section \ref{sec:recipr-cons} presents reciprocal consistency, Sections \ref{sec:distr-graph-datab}-\ref{sec:db-corruption} discuss the architecture of distributed graph databases and describes how corruption occurs. Section \ref{sec:tdelta-protocol} outlines the \tDelta protocol. Sections \ref{sec:modeling}-\ref{sec:evaluation} presents the model used to evaluate protocol performance.

>>>>

In practice, graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node graph database and motivating the need for distributed graph databases.
A common distributed graph database design pattern is to first partition graph data over several machines in a cluster; resulting in a number of \emph{distributed edges}, an edge's reciprocal pointers reside in different partitions.
Recent work \cite{Ezhilchelvan2018} and \cite{Ezhilchelvan2019} highlighted that violations of reciprocal consistency in distributed edges introduce corruption into the database.
Moreover, due to the \emph{Scale-Free} \cite{ScaleFree} property exhibited by many real world graphs, this corruption can propagate through the database at alarmingly rates.

When, for example, a BASE database \cite{Pritchett2008} is adapted with a graph processing layer, then violations of reciprocal consistency will occur if that adaptation provides no concurrency control for operations that span partitions in order to offer higher performance.
This paper proposes a simple concurrency control protocol, called the \tDelta protocol, that does not impede performance adversely.
That is because the protocol is exclusively designed for one purpose only: reciprocal consistency in distributed edges.
Its design leverages the fact that the updating of end pointers of a distributed edge \emph{must} immediately follow each other and the small interval between them is the sole \emph{raison d'\^etre} for reciprocal concurrency violations.

The paper is organized as follows: Section \ref{sec:recipr-cons} describes how edges in a graph database are stored and the notion of reciprocal consistency.
This notion is examined in the context of distributed graph databases in Section \ref{sec:distr-graph-datab} and the root causes of consistency violations are explained in the following section.
Section \ref{sec:tdelta-protocol} presents the \tDelta protocol together with correctness reasoning.
Sections  \ref{sec:modeling} and \ref{sec:evaluation} are devoted to evaluating the protocol using two metrics that measuring the protocol's ability to avoid inconsistencies (ensuring \emph{safety}) and unnecessary aborts (enhancing \emph{throughput}).
Section  \ref{sec:modeling} explains the strategies used for evaluating these metrics and the latter presents the results for various values chosen for the protocol parameter ($\Delta$). Section \ref{sec:Conclude} concludes the paper.

### 2020-04-10 paper/sec2-reciprocal-consistency.tex
camera-ready

In the property graph data model edges have direction, each edge having a pair of \emph{source} and \emph{destination} vertices.
In the storage layer, edge information is stored with \textbf{both} the source and destination vertices.
This, as we explain next, allows an edge to be traversed in both directions and thus facilitates improved query performance.

Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit.
It is expressed using vertices \emph{a} and \emph{b}, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination). The logical representation of the edge is given in Figure \ref{rc-edge}.

Using openCypher \cite{openCypher} this can be represented by:
\input{cypher/create-edge}

A common internal representation of this graph (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is given in Figure \ref{db-rep}.
A vertex is represented by a record that contains one or more vertex properties and a sequence of edge pointers pointing to other vertices.
This sequence of edge pointers is referred to as an adjacency list.

From Figure \ref{db-rep} it can be seen adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}'.
Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction.
When the adjacency list entries for a given edge are mutually compatible like this, that edge is said to be \emph{reciprocally consistent}, a form of referential integrity.
A query can read either the source or destination vertex and is able to reify the edge correctly, returning consistent results.

>>>>

In the property graph data model, edges have direction and each edge runs from a \emph{source} vertex to a \emph{destination} vertex.
In the storage layer, however, edge directionality does not exist;  \textbf{both} the source and the destination vertices store information about each other.
This allows edge traversal to be bidirectional and speeds up query performance.

Consider, for example, the statement: ``Tolkien \textit{wrote} The Hobbit''.
It is expressed using vertex  \emph{a} for Tolkien and vertex \emph{b} for The Hobbit, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination). Corresponding openCypher \cite{openCypher} code is given below and Fig \ref{rc-edge} shows the model level view.% of edge $ab$.
\input{cypher/create-edge}

Fig \ref{db-rep} depicts the internal representation of a graph arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}.
A vertex, such as \emph{a}, is represented by a record that contains one or more properties of that vertex, followed by a sequence of \emph{edge pointers} pointing to all those vertices to which this vertex is related either as a source or a destination.
The sequence of edge pointers is also called the \emph{adjacency list}.

It can be seen in Fig \ref{db-rep} that \emph{a}'s adjacency list has an edge pointer entry that stores `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s list has a corresponding entry storing the reciprocal (or inverse) information `\emph{b} \emph{written} by \emph{a}'.
When the adjacency list entries for a given edge refer to each other in a complementary manner like this, that edge is said to be \emph{reciprocally consistent}.

Consider a query: `list all titles by the author who wrote The Hobbit'.
This query needs to start from \emph{b} which represents the only entity specified explicitly in it.
Thanks to the reciprocal information in \emph{b}, it can reach \emph{a} from \emph{b}, even though edge \emph{ab} is  ``directed'' from \emph{a} to \emph{b}, and then compile the necessary list from \emph{a}.
Note that reciprocal consistency is assumed to be prevail when a query reads only the source or destination vertex of an edge.

### 2020-04-10 paper/sec3-dist-graphdb.tex
camera-ready

A distributed graph database employs a shared-nothing architecture, partitioning a graph between a number of loosely cooperating servers.
Graph partitioning is non-trivial and a common approach is to use a $k$-balanced edge cut \cite{Huang2016}.
The objective of such an approach is to minimize the proportion of edges that span partitions in a manner that balances the distribution of vertices to partitions.
Figure \ref{dist-graph} depicts a graph database partitioned across 3 servers, $S1$, $S2$ and $S3$.
Intra-partition edges are referred to as \textit{local edges} and inter-partition edges are referred to as \textit{distributed edges} (shown with dashed lines in Figure \ref{dist-graph}).
The proportion of distributed edges is always non-negligible ranging from 25-75\% \cite{Huang2016}.

@@

Adjacency lists can now contain edge pointers to vertices on remote servers. Maintaining reciprocal consistency for distributed edges is challenging - especially given an architecture employed by contemporary distributed graph databases (\cite{janusgraph}, \cite{TitanDB}). Often an existing BASE database is used for storage, which is then adapted with a query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database - we refer to such systems as BASE distributed graph databases\footnote{Typically, each partition is replicated for fault tolerance and availability, these issues are beyond the scope of this paper.}. Superficially, opting for this design appears to be a good choice: the application programmer has the modeling convenience of graphs with the operational characteristics from the underlying BASE database.

However, the problem with this design is the (lack of) transactional semantics is inherited from the underlying database. BASE databases seldom provide guarantees for multi-operation, multi-object transactions that span partitions. This lack of concurrency control across partitions makes it possible for concurrent updates to interleave in a manner that violates reciprocal consistency of distributed edges\footnote{When operating without concurrency control it is equally possible that local edges become reciprocal consistent, however primitives provided by BASE databases are typically sufficient to ensure reciprocal consistency for local edges.}. Earlier work investigated how the lack of concurrency control across partitions can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}). This process is explained in Section \ref{sec:db-corruption}.

@@

\input{figures/half-corrupted-edge}

>>>>

A distributed graph database employs a shared-nothing architecture, partitioning a graph among loosely cooperating servers.
Graph partitioning is non-trivial and a common approach is to use a $k$-balanced edge cut \cite{Huang2016}.
The objective of such an approach is to minimize the proportion of edges that span partitions and also to balance the distribution of vertices to partitions.
Fig \ref{dist-graph} depicts a graph database partitioned across 3 servers, $S_i, i = 1, 2$ and $3$.\footnote{Typically, each partition would be replicated for fault tolerance and availability, but these issues are beyond the scope of this paper.}

Intra-partition and inter-partition edges are respectively referred to as \textit{local edges}  and  \textit{distributed edges} (shown using dashed lines in Fig \ref{dist-graph}).
The proportion of distributed edges is not negligible and can range from 25-75\% \cite{Huang2016}.

@@

Adjacency lists can now contain edge pointers to vertices on remote servers.
Guaranteeing reciprocal consistency for distributed edges is challenging, especially when an existing BASE database is used for storage and is then adapted with a query language expressed in terms of edges and vertices along with some gluecode to bind that interface to the underlying database.
Opting for this design appears to be a good choice: it offers the application programmer the modeling convenience of graphs together with the operational characteristics of the underlying BASE database.

However, a major problem with this design option is the lack of transactional semantics from the underlying BASE databases.
The latter seldom provide guarantees for multi-operation, multi-object transactions that span partitions. Without concurrency control across partitions, concurrent updates of distributed edges can interleave and violate reciprocal consistency.\footnote{The concurrency control primitives provided by BASE databases are typically sufficient to ensure reciprocal consistency for local edges.}
Our earlier work showed that undermining reciprocal consistency in distributed edges can lead to irreversible corruption of the distributed database itself and the corruption rate can be worryingly high (\cite{Ezhilchelvan2018}, \cite{Webber2019}).
Origins of reciprocal inconsistency in distributed edges are explained in Section \ref{sec:db-corruption}.

@@
